Probabilistic Catalogs for Crowded Stellar Fields

### 2012-11-20 paper/ms.tex
More writing and editing

\section{Bayesian Inference}
To quantitatively model uncertainties and transform noise in observed data
into uncertainties in parameters of interest, Bayesian Inference is the appropriate
framework \citep{cox, jaynes, caticha}. Suppose there exist unknown parameters
(denoted
collectively by $\theta$) and we expect to obtain some data $x$. Our prior
state of knowledge about the parameters is modelled by a prior
probability distribution:
\begin{equation}
p(\theta).
\end{equation}
Note that this is a very concise notation \citep{hogg} and should be read
as ``the probability distribution for $\theta$''.
We also model how the parameters give rise to the data, via a generative model.
This is also known as a {\it sampling distribution}:
\begin{equation}
p(x|\theta).
\end{equation}
Despite the singular, the sampling distribution is actually a family of
probability distributions over the space of possible data sets, one probability
distribution for each possible value
of $\theta$. Note that the choice of
the sampling distribution is also an assumption about prior knowledge:
It models prior information about the fact that the data $x$ is connected to
the parameters $\theta$ in some way. Without this prior knowledge, learning is
impossible: there has to be some relationship between the parameters and the
data, otherwise new data tells you about nothing but itself.

@@

Each catalog in the posterior sample represents a scenario for the true
underlying image that we would observe if we had a hypothetical noise-free,
infinite resolution telescope. From these samples, we can construct the
posterior expected true scene. This is shown in Figure~\ref{fig:summaries} along
with the blurred (but still noise-free) version and the residuals.

@@

For Test Case 2, we then took the catalog produced by \sex~for the $0.5\sigma$
threshold and fitted a broken power-law model to the list of fluxes in the catalog.
The results of this inference are shown in Figure~\ref{fig:sex_inference}.
The inference spuriously rules out the true values of the parameters. This
occurs because the uncertainty in the stellar fluxes is not propagated through
to the inference of the luminosity function.

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{Figures/sex_inference.eps}
\end{center}
\caption{Inference of the luminosity function parameters using the $0.5\sigma$
catalog from \sex.\label{fig:sex_inference}}
\end{figure}

@@

@@

>>>>

\section{Bayesian Inference}
To quantitatively model uncertainties and transform noise in observed data
into uncertainties in parameters of interest, Bayesian Inference is the appropriate
framework \citep{cox, jaynes, caticha, mackay}. Suppose there exist unknown parameters
(denoted
collectively by $\theta$) and we expect to obtain some data $x$. Our prior
state of knowledge about the parameters is modelled by a prior
probability distribution:
\begin{equation}
p(\theta).
\end{equation}
Note that this is a very concise notation \citep{hogg} and should be read
as ``the probability distribution for $\theta$''.
We also model how the parameters give rise to the data, via a generative model.
This is also known as a {\it sampling distribution}:
\begin{equation}
p(x|\theta).
\end{equation}
Despite the singular, the sampling distribution is actually a family of
probability distributions over the space of possible data sets, one probability
distribution for each possible value
of $\theta$. Note that the choice of
the sampling distribution is also an assumption about prior knowledge:
It models prior information about the fact that the data $x$ is connected to
the parameters $\theta$ in some way. Without this prior knowledge, learning is
impossible: there has to be some relationship between the parameters and the
data, otherwise new data tells you about nothing but itself.

@@

Each catalog in the posterior sample represents a scenario for the true
underlying image that we would observe if we had a hypothetical noise-free,
infinite resolution telescope. From these samples, we can construct the
posterior expected true scene. This is shown in Figure~\ref{fig:summaries} along
with the blurred (but still noise-free) version and the residuals. The residuals
provide a check on the validity of the model, and the posterior expected true
scene provides a useful visual guide to the uncertainties present in the catalogs.

@@

For Test Case 2, we then took the catalog produced by \sex~for the $0.5\sigma$
threshold and fitted a broken power-law model to the list of fluxes in the catalog.
Such an approach might be used to study the luminosity function of a set of stars
in a field. The results of this inference are shown in Figure~\ref{fig:sex_inference}.
The inference spuriously rules out the true values of the parameters. This
occurs because the uncertainty in the stellar fluxes is not propagated through
to the inference of the luminosity function.

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{Figures/sex_inference.eps}
\end{center}
\caption{Inference of the luminosity function parameters using the $0.5\sigma$
catalog from \sex. Note that the inference appears to rule out regions around
the true input values. This occurs because the uncertainties in the existence
and fluxes of the faint stars are not taken into account.
\label{fig:sex_inference}}
\end{figure}

@@

We note that there are many limitations to the model presented in this paper.
One such limitation is that we have not considered multi-epoch or multi-band
imaging. In the former case, stellar motions may be relevant \citep{lang}, and
in the latter, variations in the spectral energy distributions of stars will
need to be considered.

@@

\bibitem[Lang et al.(2009)]{lang} Lang, D., Hogg, D.~W., 
Jester, S., \& Rix, H.-W.\ 2009, \aj, 137, 4400
